### 2011-05-12 paper/paper.tex
reworded stuff in section 3
fixed up section 5 a bit

Since the nonadaptive methods converge to the stationary distribution from any starting distribution, this yields a method that converges to the stationary distribution, but first attempts to both find a good step size in order to converge more rapidly and move to a better starting point (i.e. not near any corners).

@@

This bound is polynomial in $n$, so in theory, it is better than the Metropolis methods. The constant, however, is very large, and so our hope is that, when tested in practice, the hit-and-run method will mix to the uniform distribution even faster than in the time given by this bound.

@@

We implemented separation oracles for a cube, a sphere, and a simplex for dimensions and tested the algorithms for dimensions $n\in\{2,3,4,7,10,100\}$.
\begin{itemize}
\item \emph{The cube.} This is simply the region $(-1,1)^n$.
\item \emph{The sphere.} This is simply the region $\{(x_1,...,x_n)\in\mathbb{R}^n : x_1^2 + x_2^2 + \cdots + x_n^2 < 1\}$.
\item \emph{The simplex.} We define our simplex in the following way. Let $e_1,...,e_n$ be points in $\mathbb R^n$ defined by $e_{i,j} = 1$ if $i=j$ and $e_{i,j} = 0$ otherwise. Let $c = (1 - \sqrt{n+1}) / n$, and then let $f = (c,c,...,c)$ also be a point in $\mathbb R^n$. Then, our simplex is just the interior of the convex hull of the $n+1$ points in $V = \{f,e_1,e_2,...,e_n\}$. It is easy to compute that the distance between any two of the points in $V$ is $\sqrt{2}$, and so this simplex is indeed a regular simplex.

@@

We want to investigate how well the algorithms work in different starting places. We use the origin $(0,0,...,0)$ as a ``good'' start. We also run tests where the starting place is ``bad'', that is, near the boundary of the region. For the sphere and the simplex, we use the point $(1 - 10^{-4}, 0, ..., 0)$, and for the cube we use $(1 - 10^{-4}, 1 - 10^{-4}, ..., 1 - 10^{-4})$. Note that on the cube and simplex, the portion of the ball inside the region at a bad start is exponentially small, while for a sphere it is nearly half of the ball, so one of the theoretical boundaries to rapid mixing of the ball walking method is not present in a sphere.

@@

Ball walking poses another challenge, which is that one must sample uniformly from a ball around it. We do this by sampling uniformly from a random direction, which can be done by sampling from a Gaussian distribution and normalizing, then sample a number $\alpha$ in $[0,1)$ uniformly and set $r = \alpha^{1/n}$. To see that this is the correct distribution, note that the probability distribution of radius is proportional to $r^{n-1}$, and therefore should be equal to $nr^{n-1}$. The chosen radius is the inverse of the cumulative distribution function for this distribution, and the cumulative distribution function is easily seen to be $r^n$.

@@

For each sampling method and for each test case, we generated $1000$ samples. In most cases, we used $5000$ iterations of our random walk's transition function in order to generate each sample. However, as noted above, the hit-and-run method is slower due to its binary search. We binary search up to a precision of $10^{-12}$, which requires approximately $40$ calls to the membership oracle for each endpoint. We found empirically that transitioning only $1200$ times caused the program to run in about as much time as the other methods on $5000$ transitions. Therefore, we also ran tests with only $1200$ transitions for the hit-and-run. We also ran Gaussian-walking and adapative Gaussian-walking at $2000$ iterations for comparison between the two.

>>>>

Since the nonadaptive methods converge to the stationary distribution from any starting distribution, this yields a method that converges to the stationary distribution, but first attempts to both find a good step size in order to converge more rapidly and move to a better starting point (i.e. not near any corners), so we hope that it will perform better than the normal Metropolis method.

@@

This bound is polynomial in $n$, so in theory, it is better than the Metropolis methods, although the constant is very large. However, since theoretical constants are often pretty weak, and since this is a worst-case bound, we hope that this method will mix to the uniform distribution even faster than in the time given by this bound.

@@

We implemented separation oracles for a cube, a sphere, and a simplex for dimensions and tested the algorithms for dimensions $n\in\{2,3,4,7,10,100\}$.
\begin{itemize}
\item \emph{The cube.} This is simply the region $\{(x_1,...,x_n) \in \mathbb{R}^n : -1 < x_i < 1\text{ for all }i\}$.
\item \emph{The sphere.} This is simply the region $\{(x_1,...,x_n)\in\mathbb{R}^n : x_1^2 + x_2^2 + \cdots + x_n^2 < 1\}$.
\item \emph{The simplex.} We define our simplex in the following way. Let $e_1,...,e_n$ be points in $\mathbb R^n$ defined by $e_{i,j} = 1$ if $i=j$ and $e_{i,j} = 0$ otherwise. Let $c = (1 - \sqrt{n+1}) / n$, and then let $f = (c,c,...,c)$ also be a point in $\mathbb R^n$. Then, our simplex is just the interior of the convex hull of the $n+1$ points in $V = \{f,e_1,e_2,...,e_n\}$. It is easy to compute that the distance between any two of the points in $V$ is $\sqrt{2}$, and so this simplex is indeed a regular simplex.

@@

We want to investigate how well the algorithms work in different starting places. We use the origin $(0,0,...,0)$ as a ``good'' start (because it is in the middle or roughly in the middle, not close to the boundary). We also run tests where the starting place is ``bad'', that is, near the boundary of the region. For the sphere and the simplex, we use the point $(1 - 10^{-4}, 0, ..., 0)$, and for the cube we use $(1 - 10^{-4}, 1 - 10^{-4}, ..., 1 - 10^{-4})$. Note that on the cube and simplex, the portion of the ball inside the region at a bad start is exponentially small, while for a sphere it is nearly half of the ball, so one of the theoretical boundaries to rapid mixing of the ball walking method is not present in a sphere.

@@

Ball walking poses another challenge, which is that one must sample uniformly from a ball around it. We do this by sampling a random direction uniformly, which can be done by sampling from a Gaussian distribution and normalizing, then sample a number $\alpha$ in $[0,1)$ uniformly and set $r = \alpha^{1/n}$. To see that this is the correct distribution, note that the probability distribution of radius is proportional to $r^{n-1}$, and therefore should be equal to $nr^{n-1}$. The chosen radius is the inverse of the cumulative distribution function for this distribution, and the cumulative distribution function is easily seen to be $r^n$.

@@

For each sampling method and for each test case, we generated $1000$ samples. In most cases, we used $5000$ iterations of our random walk's transition function in order to generate each sample. However, as noted above, the hit-and-run method is slower due to its binary search. We binary search up to a precision of $10^{-12}$, which requires approximately $40$ calls to the membership oracle for each endpoint. We found empirically that transitioning only $1200$ times caused the program to run in about as much time as the other methods on $5000$ transitions. Therefore, we also ran tests with only $1200$ transitions for the hit-and-run. We also ran Gaussian-walking and adapative Gaussian-walking at $2000$ iterations, as we found this number of iterations to be good for comparison between the two.
